Reducing iptables configuration complexity using chains

### 2014-03-28 iptables_newline2014/presentation.tex
add slides for webserver and ssh/webmin only accessible for Admin PCs

      \item ssh (port 20) and webmin (port 10000), is limited to admin PCs.

@@

    \frametitle{Summary}

@@

      \item use ESTABLISHED state to reduce number of rules
      \item using chains makes your rules easier readable and maintainable
      \item chains can be reused for several rules
      \item chains can be chained together
      \item faster, because it only jumps to a chain when a rule matches

@@

   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?\\
    Contact

@@

      \item Blog : \href{http://ruleant.blogspot.com/}{http://ruleant.blogspot.com}
      \item Twitter : \href{https://twitter.com/dcadriaenssens}{@dcadriaenssens}

@@

\end{document}

>>>>

      \item SSH (port 22) and a webbased admin tool (port 10000), is limited to admin PCs.

@@

    \frametitle{Realworld example : webserver}
    Webservice, tcp port 80 (http) and 443 (https) is available for all :
      \small{iptables -A INPUT -m tcp -p tcp --dport 80 -j ACCEPT\\
    \frametitle{Realworld example : sysadmin PCs}
    Sysadmins are allowed to access SSH (tcp port 22) and a webbased admin tool (tcp port 10000).\\
    The IP addresses of their PCs :

@@

	\item 10.100.2.3
	\item 10.100.2.4
	\item 10.100.2.7

@@

    \frametitle{Realworld example : sysadmin PCs}
      \small{iptables -A INPUT -p tcp -m tcp -s 10.100.2.3 --dport 22 -j ACCEPT\\
      iptables -A INPUT -p tcp -m tcp -s 10.100.2.4 --dport 22 -j ACCEPT\\
      iptables -A INPUT -p tcp -m tcp -s 10.100.2.7 --dport 22 -j ACCEPT\\
      iptables -A INPUT -p tcp -m tcp -s 10.100.2.3 --dport 10000 -j ACCEPT\\
      iptables -A INPUT -p tcp -m tcp -s 10.100.2.4 --dport 10000 -j ACCEPT\\
      iptables -A INPUT -p tcp -m tcp -s 10.100.2.7 --dport 10000 -j ACCEPT}
    The same IP addresses are repeated. If there is a change, several rules have to be updated.\\
    \textbf{This can be done more efficient!}
    \frametitle{Realworld example : sysadmin PCs}
    Wouldn't it be convenient to :

@@

      \item make a list of IP addresses?
      \item check this list when a packet matches a rule (fe. TCP port 22)?
      \item reuse this list when another rule matches a packet?
    \textbf{This is possible with chains!}
    \frametitle{Introducing chains}
    Firstly, create the chain and add the rules
      \small{iptables -N admin\_IP\\
      iptables -A admin\_IP -s 10.100.2.3 -j ACCEPT\\
      iptables -A admin\_IP -s 10.100.2.4 -j ACCEPT\\
      iptables -A admin\_IP -s 10.100.2.7 -j ACCEPT\\
      iptables -A admin\_IP -j DROP}
    \begin{itemize}[<+->]
      \item Syntax of adding rules to a custom chain is similar to adding to default targets.
      \item Custom chains don't have a default target, so set a target for all packets that are not matched.

@@

    \frametitle{Realworld example : sysadmin PCs}
    Add the rules for TCP port 22 and 10000 using the custom chain :
      \small{iptables -A INPUT -p tcp -m tcp --dport 22 -j admin\_IP\\
      iptables -A INPUT -p tcp -m tcp --dport 10000 -j admin\_IP}

  \begin{frame}
    \frametitle{Summary}
    \begin{itemize}
      \item use ESTABLISHED state to reduce number of rules
      \item using chains makes your rules easier readable and maintainable
      \item chains can be reused for several rules
      \item chains can be chained together
      \item faster, because it only jumps to a chain when a rule matches
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?\\
    Contact
    \begin{itemize}
      \item Blog : \href{http://ruleant.blogspot.com/}{http://ruleant.blogspot.com}
      \item Twitter : \href{https://twitter.com/dcadriaenssens}{@dcadriaenssens}
    \end{itemize}
  \end{frame}
\end{document}
